---
title: "Preprocessing of OSM Building Data (HDB)"
description: "Preprocesses OSM Building Data with HDB Property Information"
author: "Teo Ren Jie"
date: "08/04/2024"
#date-modified: "2/26/2024"
number-sections: true
categories: ["Preprocessing", "Urban", "HDB", "Building", "R"]
title-block-banner: true
execute:
  message: false
  warning: false
---

# Overview

## Background

XX

# Getting Started

## Setting Up

Packages required to be loaded for

```{r}
pacman::p_load(dplyr, readr, sf, tidyverse, tmap, sfdep, ggplot2, plotly, spdep, rjson, osmextract, geojsonR, rvest, httr, jsonlite, onemapsgapi)
```

## Data Sources

| Dataset Name                                          | Source                                                                                                                                                      | Methodology |
|---------------------|---------------------------------|------------------|
| HDB Existing Buildings                                | [[Data.gov.sg]{.underline}](https://beta.data.gov.sg/datasets?resultId=d_16b157c52ed637edd6ba1232e026258d&agencies=Housing+and+Development+Board+%28HDB%29) | Download    |
| HDB Property Information (4 Aug 2024)                 | [Data.gov.sg](https://beta.data.gov.sg/datasets?resultId=d_17f5382f26140b1fdae0ba2ef6239d2f&agencies=Housing+and+Development+Board+%28HDB%29)               | Download    |
| URA Master Plan 2019 Land Use (3 Jul 2023)            | [Data.gov.sg](https://beta.data.gov.sg/datasets?query=master+plan+land+use&resultId=d_f8adc7bd980dd15861c83ff370a72eaa)                                     | Download    |
| Humanitarian OSM Team Buildings Dataset (1 July 2024) | [HOTOSM](https://data.humdata.org/dataset/hotosm_sgp_buildings?)                                                                                            | Download    |

# Data Preparation

## Loading Data

Loading HDB Property Information (4 Aug 2024)

```{r}
RAW_HDB_PROPERTY <- read.csv("data/fyp_preprocessing/HDBPropertyInformation.csv")
```

Loading HDB Existing Buildings KML

<https://stackoverflow.com/questions/50775357/how-to-read-in-kml-file-properly-in-r-or-separate-out-lumped-variables-into-col>

```{r}
raw_hdb_bldgs = st_read("data/fyp_preprocessing/HDBExistingBuilding.geojson")

attributes <- lapply(X = 1:nrow(raw_hdb_bldgs), 
                     FUN = function(x) {

                       raw_hdb_bldgs %>% 
                         slice(x) %>%
                         pull(Description) %>%
                         read_html() %>%
                         html_node("table") %>%
                         html_table(header = TRUE, trim = TRUE, dec = ".", fill = TRUE) %>%
                         as_tibble(.name_repair = ~ make.names(c("Attribute", "Value"))) %>% 
                         pivot_wider(names_from = Attribute, values_from = Value)

                     })

hdb_bldgs <- 
  raw_hdb_bldgs %>%
  bind_cols(bind_rows(attributes)) %>%
  select(-Description)
```

Loading Humanitarian OSM Buildings Dataset (1 July 2024)

```{r}
hotosm_bldgs <- st_read(dsn = "data/fyp_preprocessing/HOTOSM_BLDGS",
                   layer = "hotosm_sgp_buildings_polygons_shp") 

#%>%
#  st_transform(crs = 3414)
```

TO-DO: EXTRACT HOTOSM DATA BASED ON MP19 LAND USE POLYGON = RESIDENTIAL (CAN SELECT HDB ONLY?)

To verify whether to use OSM or HOTOSM data for buildings

Plotting HDB Existing Buildings dataset

```{r}
tmap_options(check.and.fix = TRUE)
tmap_mode("plot")
tm_shape(hdb_bldgs) +
  tm_fill(col = "brown")

```

Adding Postal Codes to HDB Property Datasets - prepare to combine with HDB Buildings geospatial data

Ref [https://renjieteo-is415.netlify.app/exercises/thex03#create-a-list-storing-unique-addressespostal-codeshttps://renjieteo-is415.netlify.app/exercises/thex03#create-a-list-storing-unique-addressespostal-codes](https://renjieteo-is415.netlify.app/exercises/thex03#create-a-list-storing-unique-addressespostal-codes)

```{r}
HDB_PROPERTY <- RAW_HDB_PROPERTY %>%
  mutate(RAW_HDB_PROPERTY, address = paste(blk_no, street))
```

```{r}
get_postal <- function(addr_lst){
  
  # Create a data frame to store all retrieved coordinates
  new_postal <- data.frame()
  i = 0
  for (addr in addr_lst){
    i = i + 1
    print(c(addr, as.character(i)))

    reply <- GET('https://www.onemap.gov.sg/api/common/elastic/search?',
           query = list(searchVal = addr,
                        returnGeom = 'N',
                        getAddrDetails = 'Y'))
    
    output <- fromJSON(rawToChar(reply$content))
    found <- output$found
    res <- output$results
    
    # Create a new data frame for each address
    new_row <- data.frame()
    
    # If single result, append 
    if (found >= 1){
      res_1 <- head(res, n = 1)
      postal <- res_1$POSTAL
      new_row <- data.frame(address = addr, postal = postal)
    }

    else {
      new_row <- data.frame(address = addr, postal = NA)
    }
    
    # Add the row
    new_postal <- rbind(new_postal, new_row)
  }
  return(new_postal)
}
```

Obtain postal codes from OneMap API

```{r eval=FALSE}
addr_lst <- sort(unique(HDB_PROPERTY$address))
postal <- get_postal(addr_lst)

write_rds(postal, "data/fyp_preprocessing/hdb_property_postal.rds")
```

Combine HDB Property data to HDB Buildings dataset

```{r eval=FALSE}
temp_hdb <- left_join(hdb_bldgs, postal, by=c("Postal" = "Postal"))
```
